---
title: "Refugees Welcome? Introducing a New Dataset on Anti-Refugee Violence in Germany, 2014--2015"
author: 
  - David Benček^[Kiel Institute for the World Economy, david.bencek@ifw-kiel.de]
  - Julia Strasheim^[GIGA German Institute of Global and Area Studies, julia.strasheim@giga-hamburg.de]
date: "March 2016"
output:
    rmarkdown::html_vignette:
        fig_caption: yes
vignette: >
  %\VignetteIndexEntry{Refugees Welcome? Introducing a New Dataset on Anti-Refugee Violence in Germany, 2014-2015}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r libraries, include=FALSE}
library(arvig)
library(lubridate)
library(plyr)
library(dplyr)
library(ggplot2)
library(ggthemes)
library(rgdal)
library(tidyr)
library(broom)
library(maptools)
```

```{r, include = FALSE, eval=FALSE}
knitr::opts_chunk$set(fig.dpi = 96)
```

# Introduction 

In 2015, an ever rising number of refugees made their journey to the European Union (EU) to seek asylum in one of its member states. 
According to the United Nations High Commissioner for Refugees (UNHCR), as of early 2016, most asylum applicants in Europe were Syrian citizens fleeing military advances by both their government as well as the Islamic State (48 per cent of arrivals), closely followed by refugees from Afghanistan (21 per cent), where a withdrawal of foreign troops has led to a resurgence of Taliban control \citep{unhcr_refugees_2016}. 
Most refugees have sought asylum in Germany and Sweden, and particularly the German reaction towards incoming refugees has sparked international attention.
By the end of summer 2015, when other member states closed their borders, Angela Merkel publicly pledged that Germany would offer temporary residence to all incoming refugees, and her government suspended applying the Dublin III Regulation that determine the member state responsible to examine asylum applications.
In addition to this "open-arms policy" \citep{hockenos_nothing_2015} of the German Chancellor, television footage of cheering citizens welcoming refugees at the Munich train station stood out in comparison to increasingly restrictive policies towards refugees across the EU.  

Not everyone welcomed refugees to Germany.
The \emph{Christlich-Soziale Union} (Christian Social Union, CSU) -- a partner in the national coalition government -- soon openly challenged Merkel's descisions, an act previously unthinkable in German consensus politics.
Simultaneously, the right-wing party \emph{Alternative für Deutschland} (Alternative for Germany, AfD) started to attract an increasing number of voters in polls.\footnote{According to one public opinion poll of February 2016, 12 percent of voters would elect the AfD into the German \emph{Bundestag} \citep{ard_ard-deutschlandtrend:_2016}.}
In addition to these political reactions, xenophobic violence directed against refugees and their supporters was on the rise in 2015 \citep{deutsche_welle_anti-refugee_2015}. 
This violence reached a tragic climax in October 2015, when Cologne city official Henriette Reker was stabbed in the neck over her position towards refugees during an electoral campaign event.

Anti-refugee violence and social unrest is not new to post-Cold War Germany, and a number of scholarly analyses have shed light on this phenomenon in the past.
To name a few examples, \citet{koopmans_discursive_2004} study the causal links between public discourse and xenophobic violence in Germany, analyzing over 11,000 public statements in the period from 1990 to 1999 \citep[cf. also][]{koopmans_explaining_1996}.
Their findings suggest that media attention to right-wing violence affects both the precise targets of such attacks as well as these attacks' temporal and spatial distribution \citep[cf. a similar analysis on the Netherlands by][]{braun_diffusion_2011}. 
\citet{krell_immigration_1996} also investigate the links between rising numbers of asylum seekers in Germany during the early 1990s and anti-refugee violence, presenting both a typology of the perpetrators as well as studying the explanatory power of various theories in order to account for the rising number of attacks.
And Willems similarly focuses on the perpetrators of right-wing violence in Germany by analyzing police data on their biographical and socio-demographic characteristics \citep{willems_development_1995} as well as public opinion polls, arguing inter alia that anti-refugee activist groups are far too heterogeneous "to be sweepingly labeled as racists" \citep{willems_rightwing_1995}.

The recent spread of anti-refugee sentiments in German politics and society has already sparked academic interest, but investigations have thus far overwhelmingly concentrated on explaining the rise of the right-wing movement Pegida (\emph{Patriotische Europäer gegen die Islamisierung des Abendlandes}, Patriotic Europeans against the Islamization of the Occicent) that flourished in late 2014 \citep[e.g.][]{dostal_pegida_2015, vorlander_pegida._2016}. 
A lack of systematic data collection and data processing of the recent anti-refugee events in Germany has thus far inhibited a thorough quantitative investigation of this phenomenon, its patterns, dynamics, drivers, and consequences. 
This paper therefore introduces a new geo-referenced event dataset on Anti-Refugee Violence and Social Unrest in Germany (hereafter ARVIG) between 2014 and 2015. 
Our dataset is based on information collected by two civil society organizations that we process to make it usable for statistical research. 
The dataset identifies in total 1\,645 events of four different types of right-wing violence and social unrest: demonstrations, assault, arson attacks, and miscellaneous attacks against refugee housing. 
In the following sections we first present data sources, discuss the categorization of different types of right-wing violence, describe the process of constructing the dataset, before outlining the variables included in the dataset. 
Afterwards we discuss initial descriptive statistics of patterns of anti-refugee violence and social unrest in Germany and conclude the paper by outlining several potential uses of the dataset in future research.


# Creating the Dataset
In order to create the ARVIG dataset, we rely on information released on the website \emph{Mut Gegen Rechte Gewalt} (MGRG), a project that was started in August 2000 by the Amadeu Antonio Foundation and the weekly magazine Stern. 
This website provides a public chronicle of anti-refugee violence and social unrest since 2014 and we include all available entries between 01.01.2014 and 31.12.2015 in the dataset.\footnote{Quarterly updates of the dataset are planned, provided that MGRG keeps publishing the information.}
The chronicle provided by the MGRG project is itself based on information collected by two civil society organizations. 
The first is the Amadeu Antonio Foundation itself that was named after Angolan citizen Amadeu Antonio Kiowa, who was one of the first victims of right-wing violence in reunified Germany when he was beaten to death by extremist youths in 1990. 
The foundation was started in 1998 with the explicit goal to strengthen German civil society activism against right-wing extremism, racism, and anti-Semitism \citep{amadeu_antonio_stiftung_amadeu_2016}.
The second organization is PRO ASYL that was founded in 1986, shortly after significant restrictions were introduced to the German asylum law that resulted in greater difficulties for people persecuted in their home countries to secure lasting protection in Germany \citep{forderverein_pro_asyl_e.v._pro_2016}.
Both the Amadeu Antonio Foundation and PRO ASYL belong to the largest and most respected pro immigration advocacy organizations and closely work together with international human rights organizations.

```{r data, echo=FALSE, cache=TRUE}
# event data
data(arvig)

# separate multi-category events into multiple rows
arvig_separated <- ddply(arvig, .(category_de), function(x){
  if (is.na(x$category_en[1])) return(x)
  splits_group <- stringi::stri_count_fixed(x$category_en[1], "&")
  if (splits_group==1) {
    two <- x %>%
      separate(category_en, c("cat1", "cat2"), sep = " & ") %>%
      gather(helper, category_en, c(cat1, cat2)) %>%
      select(date, location, state, community_id, longitude, latitude, 
             category_de, category_en, description, `source`)
    return(two)
  }
  if (splits_group==2) {
    three <- x %>%
      separate(category_en, c("cat1", "cat2", "cat3"), sep = " & ") %>%
      gather(helper, category_en, c(cat1, cat2, cat3)) %>%
      select(date, location, state, community_id, longitude, latitude, 
             category_de, category_en, description, `source`)
    return(three)
  }
  x
}) %>%
  arrange(date)
```

## Categories of Right-Wing Violence

The chronicle provided by the MGRG project documents four different types of attacks and unrest against refugees and refugee housing in Germany: demonstrations, assault, arson attacks, and miscellaneous attacks against refugee housing.
The collection is based on public reporting in newspaper articles, press releases by the German police, parliamentary interpellations as well as publicly accessible reports by local and regional organizations offering advice and consultation for victims of right-wing violence \citep{amadeu_antonio_stiftung_mut_2016}.

The first type of violence and social unrest reported by MGRG are events of anti-refugee demonstrations, such as the  rallies staged by Pegida since December 2014.
To give one example of the demonstrations the project reports, on 14 March 2015, 180 people protested against the construction of a new refugee shelter in the city of Flöha in Saxony.
The demonstration was registered by Pegida-spokesperson Steffen Musolt and at least one man was reported shouting "Sieg Heil!" \citep{freie_presse_180_2015}.
Notably, MGRG points out that because anti-refugee demonstrations and rallies have been on the rise in recent years, it is impossible to collect information on every single one of them. 
Thus, demonstrations can be expected to be under-reported in the chronicle -- and thus also in the dataset presented here.\footnote{MGRG also notes that under-reporting has become a problem mostly since January 2016, at which point the Amadeu Antonio Foundation and PRO ASYL have limited themselves to reporting demonstrations that specifically disregarded German law, meaning that the demonstration was illegal and not registered with the authorities beforehand, the demonstration included assaults against journalists or police, or demonstrators were reported using hate speech (\emph{Volksverhetzung}) \citep{amadeu_antonio_stiftung_mut_2016}. 
Since January 2016 is outside the scope of this dataset, we do not expect severe under-reporting for our observation period.}

The second type of violence reported by the MGRG project concerns physical assaults and bodily injuries.
For instance, on  12 January 2015, a Libyan asylum seeker was heavily injured in Dresden.
He had been asked for cigarettes by "men wearing bomber jackets," and after he did not understand the question, one of the men reportedly poured hot liquid over his face, shoulders, and arms, making him seek medical treatment \citep{morgenpost_weiterer_2015}.
There exist some limitations to the reporting of this type of violence as well: 
MGRG notes that information on assaults is only recorded in the chronicle if assault is carried out against an individual with refugee status.
Assaults against, for instance, left-wing protesters, volunteers helping incoming refugees, or journalists covering xenophobic rallies, are not recorded.
MGRG furthermore points out that the actual number of assaults -- independent of the status of the victim -- is likely to be much higher than what is reported in the chronicle \citep{amadeu_antonio_stiftung_mut_2016}.

The third and fourth type of anti-refugee violence reported by MGRG provides information on arson attacks against refugee housing, as well as on miscellaneous attacks against such shelters. 
For instance, on 23 March 2015, a group of unknown attackers was reported trying to set fire to a school in Berlin-Kreuzberg that houses refugees, an attack during which luckily noone was injured \citep{berlin_online_brandanschlag_2015}. 
Miscellaneous attacks against refugee housing comprise instances of rocks thrown at shelters or xenophobic graffiti.
For example, on 08 January 2015, unknown attackers painted swastikas on the walls of a house in Hausberge/Porta Westfalica (North Rhine-Westphalia) that was supposed to be turned into a refugee shelter \citep{mindener_tageblatt_haus_2015}.

In addition to these four distinct categories, some of the reported events include mixed forms of anti-refugee attacks, such as demonstrations at which refugee shelters were attacked:
On 06 March 2015, an anti-asylum demonstration of 1500 people in Freital (Saxony) not only attacked police officers and journalists with pyrotechnics, but some demonstrators also forced their way to a refugee shelter and reportedly vandalized the building (Tagesspiegel, 2015). 
Table \ref{tab:eventtypes} summarizes the frequencies of all observed event types as reported by the MGRG project.
Figure \ref{fig:event_map} offers a geographic overview of all recorded events.
For a clearer presentation, multi-category events have been split and mapped to each base category.

```{r event_frequencies, echo=FALSE}
frequency_table <- arvig %>%
  count(category_en) %>%
  rename(Category = category_en, 
         N = n)
frequency_table[is.na(frequency_table$Category),]$Category <- "other"
knitr::kable(frequency_table, caption = "Frequencies of Event Categories\\label{tab:eventtypes}")
```


By including events from this broad set of categories, our dataset covers a wider range of anti-refugee violence than some previous studies on the topic.
For instance, studying right-wing violence against asylum seekers in the Netherlands, \citet{braun_diffusion_2011} relies on data on the timing and location of events provided by the \emph{Anne Frank Stichting}, which defines right-wing violence as "[purposive] infliction of material or physical damage to targets, chosen because of their different cultural, national, ethnic, racial or religious background" \citep[cited in][]{braun_diffusion_2011}.
In our categorization, Braun's conceptualization of anti-refugee violence would thus only cover the "assault" category. 
We however prefer our broad conceptualization of anti-refugee violence and social unrest for two reasons:
First, a broad concept that also includes demonstrations and attacks against property allows researchers not only to distinguish between different types of xenophobic attacks, but also to study degrees of severity.
In that regard, our dataset can be used to help answer research questions that deal with the escalation of xenophobic extremism over time (cf. section \ref{sec:research}). 
Second, our broad conceptualization also relates more closely to other recent event data collections on global instances of violence and social unrest, such as the Social Conflict Analysis Database (SCAD) that covers protests, riots, strikes, inter-communal conflict, and government violence against civilians in Africa and Central America \citep{salehyan_social_2012}.


```{r events_map, echo=FALSE, fig.height=9, fig.width=7, fig.cap="Geographic Overview of Events by Category\\label{fig:event_map}"}
## shapefile
load("./germany.Rdata")
germany_df <- tidy(germany, region="RS")

# plot events on map
events_plot <- ggplot(germany_df) + 
  geom_map(map = germany_df, aes(long, lat, map_id = id), 
           colour = "#f2f2f2", fill ="#d9d9d9", size = 0.1) + 
  geom_point(data = arvig_separated, 
             aes(longitude, latitude, colour = factor(category_en)), alpha=0.6, size=3) +
  scale_color_colorblind(name="") +
  guides(colour = guide_legend(ncol = 2, override.aes = list(size=5))) +
  coord_map("vandergrinten") + 
  theme_map() +
  theme(legend.position="bottom", legend.key = element_blank(),
        legend.text=element_text(size=12), #legend.key.size = unit(1, "cm"),
        legend.title=element_text(size=12))

plot(events_plot)
```


## Webscraping and Geocoding

To construct the ARVIG dataset, we primarily relied on webscraping the information from the MGRG chronicle.
This is possible for all events from 01 January 2015 onwards as they are neatly separated in the HTML code of the MGRG website, using the rvest package in R that was designed in 2014 to harvest data from HTML web pages \citep{wickham_rvest:_2015}.
For the 2014 events, webscraping proved insufficient, because the entries on the MGRG website are not as neatly structured in the HTML code. 
Hence, we manually copied the 2014 events, cleaned the data and merged it with the 2015 events.

Next, we extracted the information on the location and the respective federal state from the dataset and used the Google Maps API to geocode the location (it proved necessary to take both location and federal state, to avoid confusion between two locations with the same name, e.g. Friedberg (Hessen) and Friedberg (Bayern)).
Each event is thereby mapped to a longitude and latitude with municipality-level precision.
This enables us to place each event on a high resolution map of Germany that includes geospatial information on all 11\,306 German municipalities (\emph{Gemeinden}) and determine the corresponding official 12-digit Community Identitification Number (\emph{Regionalschlüssel}).\footnote{The partition of municipalities in Germany is constantly changing -- our dataset classifies the events based on the status of 01 January 2015.}

## Variables and Patterns

```{r day_counts, echo=FALSE}
day_counts <- arvig_separated %>%
  count(date) %>%
  arrange(desc(n))
```

The ARVIG dataset contains 10 variables that characterize each recorded event. 
First, we provide the exact \emph{date} at which an event of interest occurred. 
Currently, all dates lie between 01 January 2014 and 31 December 2015. 
Events were recorded on `r nrow(day_counts)` of the 730 days covered by the dataset (cf. Figure \ref{fig:histogram}).
While on 173 days in 2014 and 2015 thus no event took place, an almost equal number of days saw at least one event of anti-refugee violence or social unrest. 
The date where most events were recorded was 29 August 2015, with a staggering count of 17 anti-refugee events, including six demonstrations, ten miscellaneous attacks against refugee shelters, as well as one instance of assault occurring in Halle (Saale), where a refugee from Guinea-Bissau was insulted, beaten and kicked by six to eight individuals. 

```{r day_histogram, echo=FALSE, fig.width = 6, fig.height = 4, fig.cap="Histogram: Events per Day\\label{fig:histogram}"}
day_histogram <- arvig_separated %>% 
  count(date) %>%
  right_join(data.frame(date=seq(dmy("01.01.2014"), dmy("31.12.2015"), by="day")), by = "date")
day_histogram$n[is.na(day_histogram$n)] <- 0

day_histogram_plot <- ggplot(day_histogram, aes(n)) +
  geom_histogram(binwidth = 1, boundary = -0.5, fill="#E69F00") +
  xlab("Events per day") +
  ylab("Days") +
  theme_bw()
print(day_histogram_plot)
```




```{r state_counts, echo=FALSE}
state_counts <- arvig_separated %>%
  mutate(state_nr = as.numeric(substr(community_id, 1, 2))) %>%
  filter(!is.na(category_en)) %>%
  count(state_nr, category_en) %>%
  arrange(desc(n))


inhabitants <- germany@data %>%
  select(RS, EWZ) %>%
  filter(EWZ > 0) %>%
  mutate(state_nr = as.numeric(substr(RS, 1, 2))) %>%
  group_by(state_nr) %>%
  summarise(inhabitants = sum(EWZ))

states <- arvig %>%
  mutate(state_nr = as.numeric(substr(community_id, 1, 2))) %>%
  dplyr::select(state, state_nr) %>%
  distinct(state, state_nr) %>%
  mutate(state = reorder(state, state_nr))

state_counts <- state_counts %>%
  left_join(inhabitants, "state_nr") %>%
  mutate(per_100k = n/inhabitants*100000) %>%
  left_join(states, "state_nr") %>%
  mutate(east = state_nr > 10) %>%
  arrange(state, category_en)
```

```{r state_counts_plot, echo=FALSE, fig.width=6, fig.height=4, fig.cap="Events per 100\\\\,000 inhabitants by state \\label{fig:state_counts}"}
state_counts_plot <- ggplot(state_counts, aes(state, per_100k, fill=category_en)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("#e69f00", "#d55e00", "#0072b2", "#009e73")) +
  xlab("") +
  ylab("Events per 100 000 inhabitants") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1),
        legend.title = element_blank(),
        legend.position=c(0.02, 0.99), 
        legend.justification=c(0,1),
        legend.key = element_rect(colour = NA),
        legend.background = element_rect(colour = "grey", size = .5))
print(state_counts_plot)
```

Next our dataset specifies the  *location* and *federal state* of events as reported by MGRG (in German writing, thus including umlauts).
All federal states of Germany have seen right-wing violence and social unrest in 2014 and 2015, but with strong variation in the number of events. 
By far the highest number of anti-refugee violence and unrest was recorded in Saxony, with `r state_counts[state_counts$state=="Sachsen",]$n` events in the 24 months under analysis, followed by North Rhine-Westphalia (`r state_counts[state_counts$state=="Nordrhein-Westfalen",]$n` events) and Berlin (`r state_counts[state_counts$state=="Berlin",]$n` events). 
The traditionally left-wing governed Bremen (where all State Premiers since 1945 have belonged to the Social Democratic Party), on the other hand, saw only `r state_counts[state_counts$state=="Bremen",]$n` events in the past two years, two arson attacks on 26 October 2015 and 26 September 2015.
If we control for state inhabitants, the densely populated North Rhine-Westphalia drops out of the top three and is replaced by Mecklenburg-Vorpommern -- a federal state that has a history of xenophobic violence against asylum seekers, for instance when between 22 and 24 August 1992, several hundred violent protesters in the Lichtenhagen district of Rostock threw stones and petrol bombs at a refugee shelter and were applauded by an even larger crowd of bystanders.
Figure \ref{fig:state_counts} depicts the number of events per 100\,000 inhabitants for all federal states and shows a clear divide between West and East Germany in the number of events directed against refugees.

Figure \ref{fig:map_district_counts} additionally depicts this relationship at the district level.
This map again highlights the high number of anti-refugee events taking place in Eastern Germany, with a particularly high count in the district of Saxon Switzerland-East Ore Mountains (\emph{Sächsische Schweiz-Osterzgebirge}). 
This district had in total 67 events of anti-refugee violence, including 10 instances of assault, 5 instances of arson, and 21 miscellaneous attacks in 2014 and 2015, while being sparsely populated with only 245\,954 inhabitants -- fewer than cities such as Mannheim, Karlsruhe, or Bonn. 


```{r intensity_map_district, echo=FALSE, fig.height=9, fig.width=7, fig.cap="Events per 100 000 inhabitants by districts\\label{fig:map_district_counts}"}
district_counts <- arvig_separated %>%
  mutate(id = substr(community_id, 1, 5)) %>%
  count(id)

district_data <- germany@data %>%
  select(RS, EWZ) %>%
  filter(EWZ > 0) %>%
  rename(id = RS,
         population = EWZ) %>%
  mutate(id = as.character(id)) %>%
  left_join(district_counts, "id") %>%
  mutate(per_100k = ifelse(is.na(n), 0, n/population * 100000),
         per_100k_f = cut(per_100k, c(0, 5, 10, 15, 20, 25,30), ordered_result = TRUE)) %>%
  right_join(germany_df, "id")

holes <- unique(district_data[district_data$hole == TRUE,]$id)

intensity_map_district <- ggplot(district_data, aes(long, lat, group = id, fill = per_100k_f)) +
  geom_map(map = district_data, aes(map_id = id)) +
  geom_map(map = filter(district_data, id %in% holes), 
           data = filter(district_data, id %in% holes),
           aes(map_id = id), colour = "#f2f2f2", size = 0.1) +
  geom_map(map = filter(district_data, !id %in% holes), 
           data = filter(district_data, !id %in% holes),
           aes(map_id = id), colour = "#f2f2f2", size = 0.1) +
  scale_fill_brewer(name="", palette = "YlOrRd", na.value="#d9d9d9") +
#   scale_fill_gradient("Events per\n100000 inhabitants", low = "#ffeda0", high = "#de2d26", 
#                        na.value="#d9d9d9") +
  coord_map("vandergrinten") + 
  theme_map() +
  theme(legend.position="bottom", legend.text=element_text(size=12), 
        legend.key.size = unit(0.75, "cm"))
print(intensity_map_district)
```


```{r municipality_counts, echo=FALSE}
municipality_counts <- arvig_separated %>%
  count(community_id) %>%
  arrange(desc(n))

berlin <- arvig_separated %>%
  filter(community_id=="110000000000") %>%
  count(category_en) %>%
  mutate(share = n/sum(n))
```

In order to facilitate disaggregated analyses of the data, the ARVIG dataset also contains the 12-digit *Community Identification Number* of the respective municipality each event has occured in.
This standardized identifier is taken from official statistics and thus makes it easy to merge highly disaggregated data from other sources with the ARVIG dataset.
<!---`Sollen wir noch den Aufbau der Community ID erklären? In einer Tabelle? JA! :)`--->
The data show that `r nrow(municipality_counts)` individual municipalities within Germany experienced right-wing extremist violence and social unrest against refugees in 2014 and 2015. 
Berlin has seen the largest number of events (`r max(municipality_counts$n)`), `r round(berlin[berlin$category_en=="demonstration" & !is.na(berlin$category_en),]$share*100, 0)`% of which were demonstrations and `r round(berlin[berlin$category_en=="miscellaneous attack" & !is.na(berlin$category_en),]$share*100, 0)`% were miscellaneous attacks.
For more detailed spatial analyses we also provide *longitude* and *latitude* of the respective event location.

The dataset furthermore contains the *event category* provided by MGRG both in German and in English.
This enables us to observe an interesting variation as to when events occurred when we analyse each type of event separately: 
For figure \ref{fig:weekdays} we split up multi-category events and added them once to each of their respective categories.
We can observe that in terms of their distribution over weekdays, assault, arson and miscellaneous attacks behave quite similarly: 
They are relatively evenly distributed over all seven days of the week, with minor spikes on the weekends (more prominent for assault and miscellaneous attacks on Saturdays).
Demonstrations, on the other hand, show a very strong spike on Saturdays.
This could indicate that while demonstrations are planned and organized -- they must be registered with the police beforehand, after all -- the other three types of anti-refugee violence and unrest occur more spontaneously. 
This distribution is at least to some extent suprising because past research has found that acts of right-wing violence in Germany occur disproportionately often on weekend nights (Willems et al. 1993), a finding that is usually linked to alcohol consumption of young men. 
Also \citet{braun_diffusion_2011} has found that weekend nights -- and summer periods -- are strong predictors for an increased hazard of xenophobic violence in the Netherlands, and (as we have explained above) Braun does not even include demonstrations in his categorization of anti-refugee violence, which seems to be the main driver of weekend occurrences in our dataset. 

```{r weekdays_plot, echo=FALSE, fig.width = 6, fig.height = 4, fig.cap="Events by Weekdays\\label{fig:weekdays}"}
## distribution over weekdays
weekdays <- arvig_separated %>% 
  filter(!is.na(category_en)) %>%
  mutate(day_of_week = wday((date), label = TRUE)) %>%
  count(category_en, day_of_week) %>%
  mutate(day_of_week = ordered(day_of_week, levels = c("Mon", "Tues", "Wed", "Thurs", "Fri", "Sat", "Sun")))

weekdays_plot <- ggplot(weekdays, aes(day_of_week, n, fill=category_en)) +
  geom_bar(stat = "identity") +
  scale_x_discrete(name="") +
  scale_y_continuous(name="Count") +
  scale_fill_colorblind(guide=FALSE) +
  facet_wrap(~ category_en) +
  theme_bw()
print(weekdays_plot)
```

Finally, a *description* of the event (in German language) as as well as a \emph{source} of this description, such as a link to a news website is provided. 
To illustrate this set of variables, table \ref{tab:sample} provides a sample record from the dataset.

```{r sample_table, echo=FALSE}
# Die LaTeX-Ausgabe muss ich noch anpassen, sodass der Zeilenumbruch funktioniert.
sample_table <- filter(arvig, date==dmy("06.03.2015"), location=="Freital") %>%
  colwise(function(x)as.character(x))(.) %>%
  gather(Variable, Sample)

knitr::kable(sample_table, caption="Sample Event from the ARVIG dataset\\label{tab:sample}")
```


# Conclusion: Using the Dataset
\label{sec:research}

This paper introduced new geo-referenced event data on anti-refugee violence and social unrest in Germany in 2014 and 2015.
Based on a public chronicle  provided by the \emph{Mut Gegen Rechte Gewalt} project, we webscraped and processed the available information so as to make them accessible for statistical research on anti-refugee violence. 
The dataset thereby complements existing research on the determinants and effects of anti-refugee attacks in Germany with new and systematic data. 
The event-based coding as well as the adherent information on event-locations make the data useful for a variety of analyses, both event-based or aggregated to German administrative units such as the communities (\emph{Gemeinden}) or districts (\emph{Kreise}). 
In that regard, the data presented here offer a starting point to analyze the recent rise in anti-refugee violence from different disciplinary backgrounds, including, but not limited to, criminology, political science, or economics. 

For instance, and as we have outlined in the introduction to this article, a number of studies have reflected upon the determinants of anti-refugee violence in the 1990s:
(youth) unemployment, the success of right-wing political parties, and media discourses have been identified as strong predictors of violent outbursts.
It would be interesting to know if the recent rise of anti-refugee attacks adheres the same old pattern, or whether new predictors have stronger explanatory power.
Similarly, our descriptive and preliminary analysis of patterns in anti-refugee violence in Germany in 2014 and 2015 already pointed to a suprising finding, namely that besides the occurrence of demonstrations, other types of anti-refugee violence are spread evenly across weekdays. 
As this goes directly against previous findings on the issue, more research on why this is the case would be advantageous. Can we detect, for instance, similar developments for other types of crime? 

Last but not least, events that have gained a lot of media attention have shaped public and political discourse.
As the impact of discourse on the outbreak of violence has been established by previous research, our event-based data enable researchers to identify key events that may have increased or decreased the ensuing level of violence.


# References
